### 2025-05-03 latex/paper/sections/08-netlist.tex
rw hello world

\section{What is a Netlist?}
\label{sec:netlist}
In its most general form, a netlist is a list of every component in an electronic design paired with a list of nets they connect to. 
Depending on the working abstraction level, these components can be transistors, logic gates, macrocells, or increasingly higher-level modules. 
Generally, a net denotes any group of two or more interconnected components.
In an electronics context, a net can be though of as a wire connecting multiple pins between multiple components, with each wire having one voltage source and one or more voltage sinks. 
Thus, if one were to express the netlist as a graph, the graph would consist of nodes (components) connected by hyperedges (wires). 
More precisely, these hyperedges connect the pins between the components, not the components themselves, with each component exposing multiple pins or ports.

In FPGA context, the components are logical cells (\texttt{LUTs}, \texttt{CARRY4s}, etc.) or hierarchical cells (Verilog module instances) with pins connected together by wires. 
In Vivado, a Netlist can be synthesized as a Hierarchical or a Flattened netlist. 
Figure ~\ref{fig:hierarchical_design} shows an example a Verilog design using module instance hierarchy, while figure ~\ref{fig:hier_netlist} shows the design synthesized into a hierarchical netlist with \textbf{hierarchical cells} and \textbf{leaf cells}. 
The synthesizer attempts to construct the module hierarchy as close to the module instantiation hierarchy defined by the user design entry. 
Figure ~\ref{fig:flat_netlist} shows the same design but synthesized into a flattened netlist.

In either synthesized netlist, the \textbf{leaf cells}, (deepest level cells), must necessarily consist only of \textbf{primitive cells} from the architecture's primitive cell library (\texttt{LUT6}, \texttt{FDRE}, \texttt{CARRY4}, \texttt{DSP48E1}, etc.). 
The netlist can be compiled and exported as a purely structural low-level Verilog file, or an Electrinic Design Interchange Format (EDIF) file, both describing the netlist explicitly as a list of logical cells connected by a list of wires.

\end{multicols}
{
    \raggedright
    \includegraphics[valign=t, scale=0.3]{figures/netlist_synth/top_level.png}
    \includegraphics[valign=t, scale=0.3]{figures/netlist_synth/module_0.png}
    \includegraphics[valign=t, scale=0.3]{figures/netlist_synth/module_1.png}
    \includegraphics[valign=t, scale=0.3]{figures/netlist_synth/module_2.png}
    \includegraphics[valign=t, scale=0.3]{figures/netlist_synth/module_3.png}
    \captionof{figure}{A simple HDL design with module hierarchy.}
    \label{fig:hierarchical_design}
}
{
    \centering
    \includegraphics[valign=c, width=11.5cm]{figures/netlist_synth/hier_netlist.png}
    \includegraphics[valign=c, width=6cm]{figures/netlist_synth/hier_graph.png}
    \captionof{figure}{
        \textbf{Left:} A hierarchical netlist consisting of LUTs and FFs.
        \textbf{Right:} The cell hierarchy graph.
    }
    \label{fig:hier_netlist}
}
\vspace{0.5cm}
{
    \centering
    \includegraphics[valign=c, width=10cm]{figures/netlist_synth/flat_netlist.png}
    \includegraphics[valign=c, width=4cm]{figures/netlist_synth/flat_graph.png}
    \captionof{figure}{
        \textbf{Left:} A flattened netlist consisting of LUTs and FFs.
        \textbf{Right:} The flattened cell hierarchy graph.
    }
    \label{fig:flat_netlist}
}
\begin{multicols}{2}

\subsection{Netlist Traversal and Manipulation in RapidWright}

One of RapidWright's most powerful features is netlist manipulation via the \texttt{edif} package of classes. A netlist can be easily extracted from a \texttt{.dcp} design checkpoint and traversed like the following:

\begin{lstlisting}[language=Java, caption={Netlist extraction and traversal}, label={lst:netlist_extract}]
Design design = Design.readCheckpoint("synth.dcp")
EDIFNetlist = design.getNetlist();

// Example task:
// Extract the set of all nets from the design.

// Initialize a new Set:
Set<EDIFHierNet> netSet = new HashSet<>();

// Access every cell:
List<EDIFHierCellInst> ehcis = EDIFNetlist.getAllLeafHierCellInstances();

// Traverse cell list:
for (EDIFHierCellInst ehci : ehcis) {
    // Access every port on this cell:
    List<EDIFHierPortInst> ehpis = ehci.getHierPortInsts();
    for (EDIFHierPortInst ehpi : ehpis) {
        // Access the net on this port:
        EDIFHierNet ehn = ehpi.getHierarchicalNet();
        netSet.add(ehn);
    }
}

// Example downstream task:
// For each net, access the source cells and sink cells.
for (EDIFHierNet ehn : netSet) {
    // Access the source ports (there should only be one)
    List<EDIFHierPortInst> sourcePorts = ehn.getLeafHierPortInsts(true, false);
    // Access the source cell
    EDIFHierCellInst sourceCell = sourcePorts.get(0).getHierarchicalInsts();

    // Access the sink ports
    List<EDIFHierPortInst> sinkPorts = ehn.getLeafHierPortInsts(false, true);
    // Access the sink cells
    List<EDIFHierCellInst> sinkCells = new ArrayList<>();
    for (EDIFHierPortInst ehpi : sinkPorts) {
        sinkCells.add(ehpi.getHierarchicalInst());
    }
}

\end{lstlisting}

The \texttt{Netlist} class in RapidWright stores hierarchical information about the \texttt{Design} and can be accessed and manipulated