### 2025-05-07 latex/paper/sections/09-packing.tex

\subsection{Site Packing}
\label{subsec:packing}